### 2019-12-31 docs/thesis-project/article.tex
Update article.tex

\section{P2P Overlays}\label{sec:p2p-overlays}
In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers. In this regard, a P2P overlay is a logical network that abstracts the physical network underneath it, typically IP-based, and where, single-hop edges represent links between pairs of peers, as depicted in Fig. \ref{fig:network_diagrams}. Before we survey related work in the field of P2P overlays, we first list some essential properties and technical jargon related to overlays and graph modeling, that helps us understand better the pros and cons of the research we are presenting in the next subsections of this thesis proposal.

\begin{figure}[ht]
    \centering
    \subfigure{\label{fig:iplayer}\includegraphics[width=0.35\textwidth, clip=true]{images/ip-layer.png}}
    \hspace{0.2\textwidth}
    \subfigure{\label{fig:overlay}\includegraphics[width=0.35\textwidth, clip=true]{images/overlay-layer.png}}
    \caption{P2P Network Diagrams: IP layer \textit{vs.} One possible overlay abstraction}
    \label{fig:network_diagrams}
\end{figure}

An overlay is said to be \textbf{\textit{connected}} if there is a path that allows any pair of peers \textit{i}, \textit{j} to communicate with each other, i.e., there is no partitioning and no peer is isolated in the system.

The \textbf{\textit{degree}} of a peer is the total number of connections established between some peer \textit{i} and other peers. It is possible to divide the degree into \textit{in-degree} and \textit{out-degree}, i.e., the number of channels a peer \textit{i} has, through which it can only receive data or send data, respectively. When all peers in an overlay have the same degree, the overlay is said to be \textbf{\textit{regular}} or \textbf{\textit{symmetric}}, but this is not usual. We can use the \textbf{\textit{degree distribution}} to evaluate the robustness of an overlay since through it; it is possible to detect \textit{weakly connected peers} and \textit{massively connected hubs}, i.e., peers with very high degree. In general, we want to minimize the existence of hubs. In P2P networks, hubs may provide the overlay with resistance against random damage and possibly speed-up lookup operations in unstructured overlays, using biased searching; however, targeted-attacks to hubs, or natural hub failures will have more significant impacts on the connectivity of the overlay, possibly leading to partitioning or halting of the system \cite{controlling_the_hubs}. Peer degrees can also allow evaluation of fair work distribution among peers of the system, e.g., in terms of processing or bandwidth.

\textbf{\textit{Network diameter}} is the number of links that compose the longest of all the shortest paths between any two peers \textit{i}, \textit{j}. Knowing the length of all shortest paths between all possible pairs of peers \textit{i}, \textit{j} one can extract the \textbf{\textit{average path length}} and hence derive how messages, on average, will be relayed on the overlay to contact a stranger. Still regarding distances, the path stretch is the ratio between the number of links a message has to traverse, between any two peers, in the underlying network and the overlay network.

\textbf{\textit{Clustering}} directly affects connectivity and network diameter; we can measure it using the \textbf{\textit{clustering coefficient}}. This coefficient represents the percentage of neighbors a peer \textit{i} has, who are also neighbors among themselves. Like peers with high degrees, high clustering coefficients indicate the increased chance of partitioning in the system in the event of massive peer failures or high degrees of \textit{churn}. Furthermore, high clustering coefficients may indicate, in gossip-based protocols, bandwidth overhead due to redundant communications in the cluster, but has the advantage of making lookups faster even when the network is experiencing heavy load, e.g., flash crowd scenarios.

Another two terms that often appear in P2P literature are the small-world and the power-law properties. The small-world property is an umbrella term for a system with a high clustering coefficient and low average path length, yet most peers are not direct neighbors of each other. The power-law property means that there is non-uniform distribution influence in the overlay, e.g., some peers have more responsibilities or perks because they have higher capacities than the rest.

>>>>

\section{P2P Overlays}\label{sec:p2p-overlays}

\begin{figure}[t]
    \centering
    \subfigure{\label{fig:iplayer}\includegraphics[width=0.35\textwidth, clip=true]{images/ip-layer.png}}
    \hspace{0.2\textwidth}
    \subfigure{\label{fig:overlay}\includegraphics[width=0.35\textwidth, clip=true]{images/overlay-layer.png}}
    \caption{P2P Network Diagrams: IP layer \textit{vs.} One possible overlay abstraction}
    \label{fig:network_diagrams}
\end{figure}

In large-scale P2P systems, it is unfeasible and even undesirable for each peer in the network to know and explicitly interact with every other node connecting it to other peers. In this regard, a P2P overlay is a logical network that abstracts the physical network underneath it, typically IP-based, and where, single-hop edges represent links between pairs of peers, as depicted in Fig. \ref{fig:network_diagrams}. We define some properties and technical jargon related to overlays and graphs that help us understand the research we survey in the next subsections. An overlay is said to be \textbf{\textit{connected}} if there is a path that allows any pair of peers $(i, j)$ to communicate with each other, i.e., there is no partitioning and no peer is isolated in the system. The \textbf{\textit{degree}} of a peer is the total number of connections established between a peer \textit{i} and other peers. We can divide the degree into \textit{in-degree} and \textit{out-degree}, i.e., the number of channels a peer \textit{i} has, through which it can only receive data or send data, respectively. When all peers have the same degree, the overlay is said to be \textbf{\textit{regular}} or \textbf{\textit{symmetric}}. We can use the \textbf{\textit{degree distribution}} to evaluate the robustness of an overlay since through it; it is possible to detect \textit{weakly connected peers} and \textit{massively connected hubs}, i.e., peers with very high degree. Hubs may provide the overlay with resistance against random damage and possibly speed-up lookup operations in unstructured overlays that use biased searching; however, hub failures will have significant impacts on the connectivity of the overlay, severely damaging performance or even halting the system \cite{webdragons, controlling_the_hubs}. Peer degrees are also indicators of fair work distribution among peers. \textbf{\textit{Network diameter}} is the number of links that compose the longest of all the shortest paths between any two peers $(i, j)$. Knowing the length of all shortest paths between all possible pairs of peers allows us to extract the \textbf{\textit{average path length}} and derive how messages, on average, will be relayed on the overlay to contact a stranger. Still regarding distances, the path stretch is the ratio between the number of links a message has to traverse, between any two peers, in the underlying network and the overlay network. \textbf{\textit{Clustering}} directly affects connectivity and network diameter; we can measure it using the \textbf{\textit{clustering coefficient}}. This coefficient represents the percentage of neighbors a peer \textit{i} has, who are also neighbors among themselves. Like peers with high degrees, high clustering coefficients indicate the increased chance of partitioning in the system in the event of massive peer failures and \textit{churn}. Furthermore, it may indicate, in gossip-based protocols, bandwidth overhead due to redundant communications in the cluster, but has the advantage of making lookups faster even when the network is experiencing heavy load, e.g., flash crowd scenarios. Another two terms that often appear in literature are the \textbf{\textit{small-world}} and the \textbf{\textit{power-law}} properties. The small-world property is an umbrella term for a system with a high clustering coefficient and low average path length, yet most peers are not direct neighbors of each other. The power-law property means that there is non-uniform distribution influence in the overlay, e.g., some peers have more responsibilities or perks because they have higher capacities than the rest.
